Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2018-01-24 stp/paper_aaai/main.tex
Adding citations, extending soa, improving experimental results section

\usepackage{colortbl}
\usepackage{color}
\usepackage{xcolor}
\definecolor{Gray}{gray}{0.96}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

% natalia
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{multirow}
\usepackage{pgfplots}
\pgfplotsset{width=10cm,compat=1.9}
\usepgfplotslibrary{colorbrewer}

@@

\subsubsection{Example} Consider the situation where we have some simple relational database tables with columns $\langle$\textit{Surname, Event, Date}$\rangle$, and 
$\langle$\textit{Company, Festival, Address}$\rangle$.
We do not \textit{a priori} know if ``Date'' is a 
date of birth of the person or the start date of the event (c.f. Fig. \ref{FIG:onto}); or whether 
``Address'' refers to the name of the place where the festival is located, or the 
email address of the company. 
Given an ontology like the one in Fig. \ref{FIG:onto}, we would like to automatically map the new data sources to  the 
ontology.\hfill$\square$

In this paper we use Machine Learning techniques (ML) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum cost Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build a graph which includes attributes (i.e. columns) from the new
source as well as ontology classes and properties.
We name it \emph{integration graph}.
It uses information gathered from the previously mapped data sources, and is further extended with information derived from the ontology.
%Firstly, we build a \emph{integration graph} which includes the 
%attributes from the new source as well as ontology classes and properties.
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
Our contributions are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates machine learning and constraint programming.
%	\npr{symbolic AI, ontological reasoning, transitive closure}
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations.
\end{itemize}

@@

A \emph{semantic model} $m$ is a directed graph with two types of nodes: 
\emph{class nodes} and \emph{data nodes}. 
We denote them as $\mathcal{C}_m$ and $\mathcal{D}_m$, respectively.
$ \mathcal{C}_m$ corresponds to classes in the ontology 
whereas $\mathcal{D}_m$ corresponds to data 
properties. 
Edges in the semantic model correspond to properties in the ontology, as 
shown in Fig.~\ref{FIG:sem} (notice the same edges appear in Fig.~\ref{FIG:onto}).
The semantic model may have several instances of the same ontology class, that is why class nodes are enumerated.

@@

\subsection{Using Patterns}

As explained in Subsection \ref{SSEC:pattern-mining}, we also use 
graph patterns in order to incentivise the solution tree 
$T^\star$ to contain subgraphs of the alignment graph that have been frequently 
seen in the training set.
To do this, we use the support of each of the obtained patterns as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

\subsection{Unmatched Attributes}
It is common that the data sources to be integrated will have columns that 
simply cannot be matched to the ontology.
This can happen when a column of a source table contains some information that 
is uninteresting to the user, or because the ontology has not been properly 
designed. 
Examples of these situations can be found in domain specific data 
\cite{Pham:semantic} or HTML tables \cite{Ritze:matching}.
In current systems, these columns are removed in a pre-processing step as a manual effort.

@@

% semantic labeling models
To illustrate that our semantic labeling model is better suited for the
matching task, 
we compare it against the state-of-the-art model \emph{DSL}~\cite{Pham:semantic} which was shown to perform 
even better than SemanticTyper. 
%\npr{not sure whether I should explain how this model works}
We use \emph{mean reciprocal rank} (MRR) to evaluate semantic labeling models.
This measure is useful to estimate how highly the true semantic label is ranked among the top $k$ suggestions.
It is calculated the following way:
$MRR = \frac{1}{n}\sum_{i=1}^{n}{\frac{1}{rank_i}},$
where $rank_i$ is the rank of the correct semantic label for the attribute $a_i$ among the top $k$ predictions.

% evaluation strategy
We perform an evaluation strategy outlined by~\cite{taheriyan2016learning}.
Let $M_j$ be the set of $j$ known semantic models.
For each data source $s_i$ in the domain we perform experiments $t-1$ times,
where $t$ is the total number of data sources in the domain and each experiment has a different number of known semantic models $M_1, M_2, \ldots, M_{t-1}$.
The case with $M_{t-1}$ known semantic models corresponds to leave-one-out validation strategy. 
For example, in the soccer domain for the source $s_1$ we run experiments 11 times using $M_1=\{m_2\}, M_2=\{m_2, m_3\}, \ldots, M_{11}=\{m_2, m_3, \ldots, m_{12}\}$
We repeat the procedure for other data sources in the domain and then average the results.
This procedure ensures that each source is at least once in the training as well as testing datasets.

@@

% karma vs serene vs serene-pats
We evaluate \serene{} and its modification \serenepats{}, which 
uses graph patterns, against the state-of-the-art system \karma{}.
All three systems have access to the same set of matches produced by our semantic labeling model.
Fig.~\ref{fig:perf} shows their performance on the museum and soccer domains.
We report average precision and recall for the systems with regard to variable 
number of semantic models in the training set.
For \serene{} and \serenepats{} we report the first solutions found by
\chuffed{}. All three systems find solutions in less than a second on average.
We use default parameters for \karma{} which were shown to yield the best 
results~\cite{taheriyan2016learning}.
As we can see, \serene{} produces on average the best semantic models in terms 
of precision while \serenepats{} generates slightly better models in terms of 
recall.
If we consider results separately per each instance and not on average, there 
are few instances where \karma{} generates more precise models, however, our 
new approaches produce much better models in terms of recall in all considered 
instances. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% sequential solution
In Fig.~\ref{fig:loo} we show how the performance of \serene{} changes 
across sequentially found solutions.
\chuffed{} successively finds better solutions in terms of Eq. \ref{EQ:obj}.
However,
that does not translate directly into more precise semantic models.
On average, we have observed that \serene{} yields very good results already 
with the first solution found, and plateaus after the third solution.

@@

The majority of approaches to solve the \relonto{} mapping problem are based on 
heuristic rules and alignment of constraints specified within relational 
schemata and ontologies.
A very comprehensive overview and comparison of existing mapping generation 
tools based on this approach is given by~\citeasnoun{Pinkel:rodi}
and~\citeasnoun{Spanos:semweb}.
As examples, there are BootOX~\cite{Jimenez:Bootox}, MIRROR~\cite{Luciano:Mirror} and  ontop~\cite{Fagin:Clio}.
Crudely speaking, these tools first apply a default direct mapping specified by the W3C.
Further, the default ontology is enriched by using explicit and implicit schema constraints.
Finally, ontology alignment techniques are applied to match the default ontology to the target ontology.
The main advantage of these systems is the ability to run in a fully automatic setting.
Our approach is complementary to the approaches from this group and at its current stage is semi-automatic.
However, constraint programming offers a convenient framework to incorporate integrity constraints specified either within relational or ontological schema as additional constraints to govern the search for the solution.
This opens an interesting direction for further research.

@@

We build upon the work of \citeasnoun{taheriyan2016learning} and use their 
ideas for the construction of the alignment graph.
The only difference is that we introduce the \emph{unknown} and \emph{root} class nodes and as many \emph{unknown} data nodes as there are attributes in the modeled data source.
These nodes serve to capture the unmatched attributes from the source.
Though we have modified \karma{} to treat these additional nodes as well, our 
approach outperforms \karma{} since we have additional constraints for these 
nodes and use an exact algorithm to solve the STP.
\citeasnoun{taheriyan2016learning} treat the matching and STP parts of \relonto{} independently and use heuristic algorithms for both.
We, on the other hand, use exact algorithms for both parts and address them within a unified CP model.

In their follow up work, \citeasnoun{Taheriyan:Leveraging} suggest using graph 
patterns to boost the performance of their system.
To this end, they had to revise their algorithm by introducing additional heuristics.
However, in our case we only had to add pattern variables and to modify the 
objective function in the \minizinc{} model.
No changes to the solver were required.
This makes our system very convenient and opens directions for validating various additional constraints.
\ignore{Though our approach beats the state-of-the-art in the conducted 
experiments, it 
is worth noting that the scalability of our approach is under question.
The main argument why exact solution for STP is feasible in our setting is that 
we map data sources to a domain-specific ontology which is expected to be 
limited in its size.}

>>>>

\usepackage{colortbl}
\usepackage{color}
\usepackage{xcolor}
\definecolor{Gray}{gray}{0.96}
\usepackage{algorithm}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

% natalia
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{multirow}
\usepackage{pgfplots}
\pgfplotsset{width=10cm,compat=1.9}
%\usepgfplotslibrary{colorbrewer}

@@

%\subsubsection{Example} 
Consider the situation where we have some simple relational database tables with columns $\langle$\textit{Surname, Event, Date}$\rangle$, and 
$\langle$\textit{Company, Festival, Address}$\rangle$.
We do not \textit{a priori} know if ``Date'' is a 
date of birth of the person or the start date of the event (c.f. Fig. \ref{FIG:onto}); or whether 
``Address'' refers to the name of the place where the festival is located, or the 
email address of the company. 
Given an ontology like the one in Fig. \ref{FIG:onto}, we would like to automatically map the new data sources to  the 
ontology.\hfill$\square$

In this paper we use Machine Learning (ML) techniques and Constraint Programming (CP) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum cost Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build a graph which includes attributes (i.e. columns) from the new
source as well as ontology classes and properties.
We name it \emph{integration graph}.
It uses information gathered from the previously mapped data sources, and is further extended with information derived from the ontology.
%Firstly, we build a \emph{integration graph} which includes the 
%attributes from the new source as well as ontology classes and properties.
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use CP to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
Our contributions are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates techniques from the areas of machine learning, knowledge representation and constraint programming.
%	\npr{symbolic AI, ontological reasoning, transitive closure}
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes in the integration graph.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations. This is achieved since CP allows introduction of additional constraints with no changes to the solver. 
\end{itemize}

@@

A \emph{semantic model} $m$ is a directed graph with two types of nodes: 
\emph{class nodes} and \emph{data nodes}. 
We denote them as $\mathcal{C}_m$ and $\mathcal{D}_m$, respectively.
$ \mathcal{C}_m$ corresponds to classes in the ontology 
whereas $\mathcal{D}_m$ corresponds to data 
properties. 
Edges in the semantic model correspond to properties in the ontology, as 
shown in Fig.~\ref{FIG:sem} (notice the same edges appear in Fig.~\ref{FIG:onto}).
The semantic model may have several instances of the same ontology class, that is why class nodes are enumerated.
Semantic models can be formalized using a standard mapping language such as R2RML~\cite{r2rml}.

@@

\textbf{Patterns}.
As explained in Subsection \ref{SSEC:pattern-mining}, we also use 
graph patterns in order to incentivise the solution tree 
$T^\star$ to contain subgraphs of the alignment graph that have been frequently 
seen in the training set.
To do this, we use the support of each of the obtained patterns as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

\textbf{Unmatched Attributes}.
It is common that the data sources to be integrated will have columns that 
simply cannot be matched to the ontology.
This can happen when a column of a source table contains some information that 
is uninteresting to the user, or because the ontology has not been properly 
designed. 
Examples of these situations can be found in domain specific data 
\cite{Pham:semantic} or HTML tables \cite{Ritze:matching}.
In current systems, these columns are removed in a pre-processing step as a manual effort.

@@

% semantic labeling models
To illustrate that our semantic labeling model is better suited for the
matching task, 
we compare it against the state-of-the-art model \emph{DSL}~\cite{Pham:semantic} which was shown to perform 
even better than SemanticTyper. 
%\npr{not sure whether I should explain how this model works}
We use \emph{mean reciprocal rank} (MRR) to evaluate semantic labeling models.
This measure is useful to estimate how highly the true semantic label is ranked among the top $k$ suggestions.
It is calculated the following way:
$MRR = \frac{1}{n}\sum_{i=1}^{n}{\frac{1}{rank_i}},$
where $rank_i$ is the rank of the correct semantic label for the attribute $a_i$ among the top $k$ predictions and $n$ is the number of attributes in the data source.

% evaluation strategy
We perform an evaluation strategy outlined by~\cite{taheriyan2016learning}.
Let $M_j$ be the set of $j$ known semantic models.
For each data source $s_i$ in the domain we perform experiments $t-1$ times,
where $t$ is the total number of data sources in the domain and each experiment has a different number of known semantic models $M_1, M_2, \ldots, M_{t-1}$.
The case with $M_{t-1}$ known semantic models corresponds to leave-one-out validation strategy. 
For example, in the soccer domain for the source $s_1$ we run experiments 11 times using $M_1=\{m_2\}, M_2=\{m_2, m_3\}, \ldots, M_{11}=\{m_2, m_3, \ldots, m_{12}\}$
We repeat the procedure for other data sources in the domain and then average the results.
This procedure ensures that each source is at least once in the training and testing datasets.

@@

% karma vs serene vs serene-pats
We evaluate \serene{} and its modification \serenepats{}, which 
uses graph patterns, against the state-of-the-art system \karma{}.
All three systems have access to the same set of matches produced by our semantic labeling model.
Fig.~\ref{fig:perf} shows their performance on the museum and soccer domains.
We report average precision and recall for the systems with regard to variable 
number of semantic models in the training set.
For \serene{} and \serenepats{} we report the first solutions found by
\chuffed{}. All three systems find solutions in less than a second on average.
We use default parameters for \karma{} which were shown to yield the best 
results~\cite{taheriyan2016learning}.
As we can see, \serene{} produces on average the best semantic models in terms 
of precision while \serenepats{} generates slightly better models in terms of 
recall.
The t-tests conducted for the obtained results confirm statistical significance of our improvements at the level $p<0.01$
%If we consider results separately per each instance and not on average, there 
%are few instances where \karma{} generates more precise models, however, our 
%new approaches produce much better models in terms of recall in all considered 
%instances. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% sequential solution
In Fig.~\ref{fig:loo} we show how the performance of \serene{} changes 
across sequentially found solutions.
\chuffed{}, which is the solver of \serene{}, successively finds better solutions in terms of Eq. \ref{EQ:obj}.
However,
that does not translate directly into more precise semantic models.
On average, we have observed that \serene{} yields very good results already 
with the first solution found, and plateaus after the third solution.
Furthermore, we have explored how the systems perform if we manually filter out the unmatched attributes.
In such scenario \serene{} also yields better solutions than \karma{}, and in many situations \serene{} can prove optimality of the found solutions.
This observation provides further evidence towards the validity of our new modeling framework.

@@

The majority of approaches to solve the \relonto{} mapping problem are based on 
heuristic rules and alignment of constraints specified within relational 
schemata and ontologies.
A very comprehensive overview and comparison of existing mapping generation 
tools based on this approach is given by~\citeasnoun{Pinkel:rodi}
and~\citeasnoun{Spanos:semweb}.
As examples, there are BootOX~\cite{Jimenez:Bootox}, MIRROR~\cite{Luciano:Mirror} and  ontop~\cite{Fagin:Clio}.
Crudely speaking, these tools first apply a default direct mapping specified by the W3C.
Further, the default ontology is enriched by using explicit and implicit schema constraints.
Finally, ontology alignment techniques are applied to match the default ontology to the target ontology.
The main advantage of these systems is the ability to run in a fully automatic setting.
Our approach is complementary to these approaches and at its current stage is semi-automatic.
However, CP offers a convenient framework to incorporate integrity constraints specified either within relational or ontological schema as additional constraints to govern the search for the solution.
This opens an interesting direction for further research.

@@

We build upon the work of \citeasnoun{taheriyan2016learning} and use their 
ideas for the construction of the alignment graph.
The difference at this step is that we introduce the \emph{unknown} and \emph{root} class nodes and as many \emph{unknown} data nodes as there are attributes in the modeled data source.
These nodes serve to capture the unmatched attributes from the source.
Though we have modified \karma{} to treat these additional nodes as well, our 
approach outperforms \karma{} since we have additional constraints for these 
nodes and use an exact algorithm to solve the STP.
\citeasnoun{taheriyan2016learning} treat the matching and STP parts of \relonto{} independently and use heuristic algorithms for both.
We, on the other hand, use exact algorithms for both parts and address them within a unified CP model.
In their follow up work, \citeasnoun{Taheriyan:Leveraging} suggest using graph 
patterns to boost the performance of their system.
To this end, they had to revise their algorithm by introducing additional heuristics.
However, in our case we only had to add pattern variables and to modify the 
objective function in the \minizinc{} model.
No changes to the solver were required.
This makes our system very convenient and opens directions for validating various additional constraints.
\ignore{Though our approach beats the state-of-the-art in the conducted 
experiments, it 
is worth noting that the scalability of our approach is under question.
The main argument why exact solution for STP is feasible in our setting is that 
we map data sources to a domain-specific ontology which is expected to be 
limited in its size.}

Another relevant problem within the Semantic Web community is ontology matching, or alignment~\cite{shvaiko2013}.
This problem is similar to schema matching since it finds semantically related entities across different ontologies.
Yet, it is different from \relonto{} since data sources in our case are not bound to any ontology.
We think that our approach might be further extended to incorporate ontology matching techniques.
